---
title: "Mouse-Meta-Analysis"
author: "Sidhant Puntambekar"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Mouse meta-analysis

This file contains the meta-analysis of the mouseAtlas.rds generated by combining reference matrices from a variety of scRNA-seq datasets from across NCBI GEO and Tabula Muris. 

First, the necessary libraries will be loaded in. They are dplyr (standard dataframe library), Seurat (create Seurat object), patchwork (ggplot library), clustifyr (clustifyr functions), tidyverse (collection of R data science packages), here (reproducibility of results), and clustree (visualizing clustering in hierarchical structure)

```r
library(dplyr)
library(Seurat)
library(patchwork)
library(clustifyr)
library(tidyverse)
library(here)
library(clustree)
```

## Load Mouse Atlas

We will now load the mouse atlas into the R data section. This will be accomplished with the here library establishing the root of the project and then 

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
